---
title: "Prediction Assignment Writeup"
author: 
date: "23 Juni 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Goal

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

# Loading packages
```{r}
library(knitr)
library(caret)
library(rpart)
library(randomForest)
library(rattle)
library(rpart.plot)


```

# Loading and preprocessing the data 

```{r}
# set the URL for the download
Train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Test  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
training <- read.csv(url(Train))
testing  <- read.csv(url(Test))

# remove variables that are NA
var_na   <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, var_na==FALSE]
testing  <- testing[, var_na==FALSE]

# remove identification variables
training <- training[, -(1:5)]
testing  <- testing[, -(1:5)]

# remove variables with near zero variance
nzv <- nearZeroVar(training)
training <- training[, -nzv]
testing  <- testing[, -nzv]


dim(training)
dim(testing)


```
# Data partition for Cross-validation 
The training data is split into two data sets, for training the model and for testing the performance of our model. The data is partitioned by the classe variable, which is the variable we will be predicting. The data is split into 70% for training and 30% for testing.

```{r}
# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]


dim(TrainSet)
dim(TestSet)
```

# Creating a Decision Tree
To predict the classe variable, a Decision Tree is build.


```{r}

modFitDecTree <- train(classe ~ ., data=TrainSet, method="rpart")
print(modFitDecTree$finalModel)
fancyRpartPlot(modFitDecTree$finalModel)

predictDecTree <- predict(modFitDecTree, newdata=TestSet)

confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
    
```
###The accuracy of the Decision Tree model is 0.564.

###Now a Gradient Boosted Tree model is generated.

# Generating a Gradient Boosted Tree model

```{r}

modFitBoostTree<- train(classe ~ ., method="gbm",data=TrainSet,verbose=FALSE)
print(modFitBoostTree)


predictBoostTree <- predict(modFitBoostTree, newdata=TestSet)

confMatBoostTree <- confusionMatrix(predictBoostTree, TestSet$classe)
confMatBoostTree

```

###The Gradient Boosted Tree model has a better accuracy : 0.9898

# Applying the Gradient Boosted Tree model to the 20 test cases

###The Gradient Boosted Tree model will be applied to the 20 test cases available in the test data: 
```{r}

predictTEST <- predict(modFitBoostTree, newdata=testing)
predictTEST
```
